---
title: "Megabummer"
output: html_document
---

### Data Science BIO 260 Final Project, Spring 2016
##### Allison Blajda, Leo Brown, and Emily Lisker

### Overview and Motivation

Social media users leave digital traces that can be leveraged to guide consumer purchasing decisions. Twitter users, in particular, provide spatial and temporal data that can be analyzed to develop a more precise understanding of users’ perspectives or sentiments about a service, company, or product. This information, in turn, can be shared with the public to allow customers to make decisions based on crowd-sourced knowledge. It may also even help the service provider understand how they can improve their services.

Based on our own personal experiences, we aim to analyze recent tweets to understand the degree of negative experience among users of the low-cost bus company, Megabus, over time. While Megabus is an inexpensive and convenient transportation option with many buses departing major cities each day, we hypothesize that some Megabus routes and departure times may be more reliable, i.e., busses have fewer breakdowns and delays, than others. Using social media to understanding the predictors of negative Megabus experience, we may be able to offer future bus riders information that will help them avoid these experiences. 

Our primary goal is to understand the determinants of negative experience on Megabus, with a focus on temporality, in order to provide Megabus riders with information on the likelihood of service interruptions based on date and time.

### Related Work: 

TO BE COMPLETED: Anything that inspired you, such as a paper, a web site, or something we discussed in class.

### Initial Questions: 

TO BE COMPLETED: What questions are you trying to answer? How did these questions evolve over the course of the project? What new questions did you consider in the course of your analysis?

### Data: 

We scraped data from the Twitter API using the TwitteR package. We cleaned the data using dplyr and the tidytext package, available here: https://github.com/juliasilge/tidytext. 

TO BE COMPLETED: describe cleaning and add any info on python

Load relevant packages and setup Twitter authorization.

```{r}
library(twitteR)
library(dplyr)
library(ggplot2)
library(ggthemes)

# setwd(dir = "/Users/eblisker/Documents/HSPH/Courses/2016 Spring/BIO 260/Final/megabummer")
# setup_twitter_oauth(api_key,api_secret,access_token,access_token_secret)
```

Search twitter

```{r}
# tweets <- searchTwitter("megabus", n = 3500, lang="en")
# tweets_df <- bind_rows(lapply(tweets, as.data.frame))
```

### Exploratory Analysis

TO BE COMPLETED: What visualizations did you use to look at your data in different ways? What are the different statistical methods you considered? Justify the decisions you made, and show any major changes to your ideas. How did you reach these conclusions?

```{r}

```

### Final Analysis 

TO BE COMPLETED: What did you learn about the data? How did you answer the questions? How can you justify your answers?
